Forecasting interrupted time series

### 2024-02-14 fits.qmd
add some texts to the introduction

Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change. Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions. In this paper, we consider the problem of forecasting during such an event, and after it has occurred.

This is a different problem from change point detection. In the situations we consider, we know that a change has occurred, and we want to forecast the future after the change. Change point detection is about identifying *when* a change has occurred.

Time series models usually assume that the data evolve in the future in a similar way to how they have evolved in the past. But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.

@@

@@

@@

>>>>

Time series forecasting models use historical data to forecast future values based on consistent patterns observed in the past. However, time series are sometimes interrupted by unusual events, that jeopardise the regularity of the time series pattern. Such interruptions may take a variety of forms including from a temporary policy change to a natural disaster; a few examples are provided below:

- Outbreaks, epidemics, and pandemics may all have a substantial influence on population behaviour, causing fluctuations and affecting time series data in a wide range of areas such as hospital admissions, pharmaceutical sales, transportation, retail sales, tourism, transportation, and traffic.

- Supply chain interruptions, such as strikes, trade disputes, or shortages, may cause inconsistencies in data, including sales, inventory levels, production, and deliveries . Firms may need to shut down for repair, overhaul, or maintenance purposes. These events force production to halt, leaving a gap in the data.

- Equipment failures in transportation: Vehicle, aeroplane, or rail malfunctions or breakdowns may result in missing or incorrect data points in transportation-related datasets .

- Cyberattacks: Cybersecurity breaches may jeopardise data integrity, resulting in incorrect readings in datasets including financial transactions, network traffic, and user activity.

The occurrence of such events presents a tremendous challenge to time series forecasting methods, limiting their capacity to automatically and accurately capture systematic information and forecast future values. Therefore, practitioners face a critical question: how to effectively handle time series data impacted by such disruptive events?  Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions. This study is motivated by numerous conversations with academics and practitioners, which demonstrate widespread concern about dealing with data affected by events like the COVID-19 pandemic.

Time series models usually assume that the data will evolve in the future in a similar way to how it has evolved in the past. But a big event (such as COVID-19) can result in a future that is different from the past, at least in the short term.

@@

<!-- Time series are sometimes interrupted by unusual events; for example, a natural disaster may occur, or a there may be a temporary policy change. Many forecasters have faced this issue recently with the COVID-19 pandemic, where historical patterns have been severely disrupted due to lockdowns and other restrictions. -->

It is important to note that in this study, we consider the problem of forecasting both during and after the event. This is a separate problem from change point detection. In the situations we investigate, we know that a change has occurred and want to forecast what will happen next. Change point detection refers to determining *when* a change occurred.

The paper proposes and evaluates multiple forecasting strategies designed to handle interruptions in time series data caused by disruptive events. These approaches include a range of models capable of capturing different types of changes, from simple level shifts to more complex alterations in seasonal patterns and trends. By comparing the performance of these models on two real data sets, the paper offers valuable insights into the effectiveness and practical implications of each approach.The paper offers practical guidance for practitioners seeking to improve the accuracy and reliability of time series forecasts in the presence of disruptive events. By discussing the advantages and disadvantages of each strategy and providing recommendations based on the findings, the paper equips practitioners with actionable insights for addressing the challenges associated with forecasting during and after disruptive events. 
Furthermore, by adhering to reproducibility principles and providing both data and R codes for the forecasting models, the research not only contributes to transparency but also promotes the generalizability of the suggested strategies to a variety of domains. This openness allows for the replication of results, ensuring the reliability of findings, and enhancing the accessibility and utility of study outputs for the larger research community.

@@

# Related work

Stock out, maintenancem sensors

@@

# Conclusion
